### 2016-05-30 sections0.03/method0.03.tex

%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%Method
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------

\section{METHOD}
\label{sec: method}
 \subsection{Self Organizing Maps}
 \label{sec: som}

 The SOM is a clustering method which reduces the dimension of the data, usually to 1D or 2D, while preserving topological features of the original dataset.
 Results of SOM contain nodes (neurons) that are arranged in a 1D or 2D arrays \citep{Kohonen98}. 
 Each node may contain one or more samples from the input data.
 The distance between the nodes represents similarity or dissimilarity of underlying samples, i.e., similar data are closer together in the array and the longer the distance between two nodes, the more dissimilar are their samples.
 A weight vector ``\boldit{W}" with the same dimension as the input data is associated with each node, which will be varied during the training process.
 This vector is the key factor in determining the position of the nodes in a map.
 \cite{Geach12} presented the application of the SOM and discussed its algorithm in detail.
 In this section we briefly discuss the algorithm of SOM, how we create our maps and a present a  test model which will help interpreting our results.

 \subsubsection{Algorithm of SOM} 
 \label{sec: algorithm}
     Assume we have a dataset which contains vectors, \boldit{V} $\in \Re^n$, and we want to map them on an S1 by S2 map. 
     We start by creating S1 $\times$ S2 empty neurons. 
     The initial arrangement of these neurons depends on the map's topology provided by the user.
     In the case of the 1D maps, since each neuron has two immediate neighbours the topology of the map does not have any effect on the final result and any topology can be chosen.
     However, in 2D maps, the shape of the neurons specifies the number of immediate neighbours for each neuron and it is up to user to choose the most suitable shape based on the data.
     In this paper, we choose hexagonal topology, which gives each neuron six neighbours, and provides more interactions between neurons.
     Then, a random weight vector, \boldit{W} $\in \Re^n$, will be assigned to each node.
     The process of creating SOM happens over a series of $N$ iterations. 
     During each iteration the weight vectors might change according to the Kohonen learning rule (equation~\ref{equ: weight adj}). 
      In each iteration, the SOM code:
     \begin{enumerate}
        \item chooses a random vector from the dataset ($V_i$).
        \item calculates the Euclidean distance for each node, $j$, as  $D_j^2= \sum_{i=0}^{i=n} (V_i - W_i)^2$, and finds a neuron with minimum $D_j$, (``$D_{j_{min}}$"). This neuron is the winner node and is called the Best Matching Unit (BMU). 
        \item  computes the radius of the neighbourhood of the BMU to find nodes within this radius. The weight vectors of these nodes will be affected in the next steps. The radius of the neighbourhood is arbitrary and can be set to be as high as half of the SOM size. It then decays exponentially over each iteration as
        \begin{equation}
            r^t_{BMU} = r^0_{BMU}e^{(-t/\tau)}
        \end{equation}
        where $\tau$ is a decay constant and is usually set to be the same as the number of iterations, $N$. $r^0_{BMU}$ and $r^t_{BMU}$ are the radii of the neighbourhood at 0th and $t$th iteration, respectively. 
        \item changes the weight vectors of the BMU and all the nodes within $r^t_{BMU}$ as:
        \begin{equation}
            \label{equ: weight adj}
            w(t+1)=w(t)+L(t) \times R(t) \times(v(t)-w(t))
        \end{equation}
        where $L(t) = L_0 e^{(-t/\tau)}$ is the learning factor, which prevents the divergence of the SOM and $R(t)=\exp(-\frac{D_j^2}{2r^t_{BMU}})$ is the influence rate. $R(t)$ determines how the weight of each node in the neighbourhood of BMU will change.
     \end{enumerate}
     These steps are then repeated $N$ times.

\subsection{Creating SOM}
\label{sec: create_som}
     In order to create SOM, we use {\tiny MATLAB} neural network toolbox~\citep[NNT,][]{matlabtolbox}.
     %%Sr160611: Last time I got the following note from the MNRAS editor:
     %"Please note that computer software/programming lan- guages must be styled in SMALL CAPITAL LETTERS, according to journal style. Please check and correct this paper accordingly."
     SOM in {\tiny NNT} can be created by {\tiny NEWSOM} or {\tiny SELFORGMAP} library, both of which work in two phases; ``ordering phase" and ``tuning phase". 
     The first phase is called the ``ordering phase". 
     This phase starts with maximum neighbourhood distance, and an initial high learning factor (usually 0.9), which is provided by the user. 
     The ordering phase continues for a requested number of iterations.
     During the iterations, the learning factor decreases to the tuning phase learning factor and the neighbourhood distance reaches that of the tuning phase as well.
     Both the learning factor and the neighbourhood of the tuning phase are set by the user. 
     The amount by which these two factors change in each iteration depends on the number of iterations.

     The second phase is called the ``tuning phase".
     During this phase, the neighbourhood distance is kept at the user-defined minimum.
     The learning factor, however, decreases gradually.
     The gradual change in the leaning factor helps fine tuning the topology results, leading to a more stable SOM. 
     To allow the fine tuning, The number of iterations in this phase must be much larger than the that of the ordering phase. %cite kohenen book?!
     We chose the number of epochs in the tuning phase to be 3 times the number of epochs in the ordering phase.

     To present our results, we use the {\tiny NNT}'s built-in plotting tool.
     This tool, which is the enhanced version of the old SOM plots, are designed to show the distance between each cluster more clearly.
     Specifically, we use two of the plots in this tool: Hits map, which shows the number of times each neuron has won (hits), and distance map, which shows the distance between those neurons.
     In the maps, the purple hexagonal shape areas represent the neurons.
     The distances in a distance map are shown by the grey cycle colours.
     The darker the colour, the larger distance between neurons.
     In the hit maps, neurons with zero hits are left empty.

    The sizes of the SOM maps are arbitrary and there are no rules regarding choosing one over the other. 
    However, \cite{Vesanto05} suggested that the total number of $5\sqrt{n}$ neurons is the most sufficient size, but users usually choose the size of the grids based on their dataset and their application of the results.

\subsection{Mock sample}

         \begin{figure}
            \begin{subfigure}[b]{0.5\textwidth}
                \centering
                \includegraphics[width=\textwidth]{images0.01/sample/sample2_dist.png}
            \end{subfigure}
            \hfill
            \begin{subfigure}[b]{0.5\textwidth}
                \includegraphics[width=\textwidth]{images0.01/sample/sample2_hits.png}
            \end{subfigure}
            \caption{SOM of the mock sample. In both plots, the axes show the position of the neurons. Hexagonal shapes represent the neurons. The top plot is a distance map. The grey cycle colours show the differences between the weights of each neuron with white being the minimum differences and black being the maximum one. The lower plot is a ``hits plot". It represents the number of samples in each neuron, where empty means zero hits. In this sample, the 27 galaxies clustered in 4 groups, containing 8, 7, 6 and 6 galaxies respectively.}
            \label{fig: sample}
        \end{figure}

 To illustrate how SOM works, we create a mock sample of 27 galaxies.
% We create a mock sample of 27 galaxies to show how this method works, illustrate the results of SOMs, and how they can be analysed.(<- I don't like this sentence)
 The mock sample contains two attributions of each galaxy: the type (type 1 or type 2), and the start formation rate (high or low). 
 We generated a SOM sized $10 \times 10$ using the same initial values mentioned in Section ~\ref{sec: create_som}.

 Fig. ~\ref{fig: sample} shows the SOM of this mock sample. 
 The upper plot is the distance map. 
 The axes show the position of the neurons in the $10 \times 10$ network.
 The lower plot in Fig.~\ref{fig: sample} shows the hit map.
 On this map, similar to the distance map, the axes show the position of the neurons and the hexagonal shapes are the neurons.
 The purple neurons with a number on them shows the number of galaxies in them.
 Colour coverage of neurons depends on the number of the hits on the sample.
 The neuron with the maximum number of hits is darkly coloured while the empty neurons are left uncoloured (white).

Using this method, as expected, we are able to divide the mock sample galaxies into 4 distinct groups: Type 1 with high SFR, type 1 with low SFR, type 2 with high SFR and type 2 with low SFR. 
The upper plot in the Fig.~\ref{fig: sample}, clearly shows this division.
In that plot, the upper half belongs to high star forming galaxies, while the lower half belongs to the low star forming galaxies.
The left half of the plot is where type 1 galaxies belong to, and type 2 galaxies reside in the right side.
Different shades of gray show the border between regions.
The lower plot in Fig.~\ref{fig: sample} shows that only 4 neurons (out of 100) are occupied. 
8 galaxies are type 1 galaxies with high SFR, 7 are type 2 galaxies with high SFR, 6 are type 1 galaxies with low SFR and the other 6 are type 2 galaxies with low SFR. 
%Since the input data are chosen to not to be complex, each entry only could get either 0 or 1 as a type of galaxy and 0 or 0.5 as an indicator of  high or low SFR, only four of the neurons were filled.
We can conclude that although each galaxy had  the chance to occupy any of the neurons in the network, because of the similarity in the values of their attributions, they remained only in four groups.
This network is considered as a train network and can be used to cluster any new dataset with similar entries.

As we show in the following sections, with data from real galaxies, we see that the number of dimensions are more than two and we never have two galaxies with exactly the same information. 
Therefore, if the network has high enough neurons, the input data would eventually separate from each other and cluster into smaller groups. 
However, if the information of the input data is very similar to each other, the number of neurons is going to be much higher than the number of input samples, to be able to separate the groups from each other. 
Therefore, it is up to the users that decide the similarity or dissimilarity between the input data based on number of neurons.