### 2018-03-25 Z_ref_reports_responses/ref_clarification_report3.tex

"By "compare to an external standard" I mean "compare performance on a labeled sample", i.e., one labeled by humans, to see how well the proposed method (and a simple nearest-neighbors baseline) can predict those classifications.

My suggestion with the 142 galaxies was that, if their types are not previously known, to have a human label them (i.e., assign which of the 12 K96 templates each one should belong to).  That would allow for an assessment of the automated classification quality.  If no human can do this, then (as stated) there is no way to assess classification accuracy of the proposed methods.”"