### 2018-03-25 sections0.09/email_to_editor.tex

Dear Mrs Williams,

Thank you for your patience with the revisions for this paper. We are committed to completing the process. To avoid further delays, we would like to ask for clarification from the referee to ensure that we understand the request in his/her most recent report.

The particular part of the report that we are unclear about are the three paragraphs which appear before "Additional minor comments and suggestions". We don't understand what is meant by "compare to its performance (on an external standard)" - as the referee says, it's not obvious why chi-square fitting should be considered the standard.

We are also confused by the final paragraph of these comments. The true labels of the 142 galaxies are unknown, which we understand to mean that there is no external reference for the classification accuracy - but this is the opposite of what the last sentence implies. Is it possible that "true labels.. are unknown" should read "true labels ..are known"?

We would be grateful for any clarification that could be offered. We very much appreciate the referee's careful attention and thoughtful comments.

Editor's Answer:
“By "compare to an external standard" I mean "compare performance on a labeled sample", i.e., one labeled by humans, to see how well the proposed method (and a simple nearest-neighbors baseline) can predict those classifications.

My suggestion with the 142 galaxies was that, if their types are not previously known, to have a human label them (i.e., assign which of the 12 K96 templates each one should belong to).  That would allow for an assessment of the automated classification quality.  If no human can do this, then (as stated) there is no way to assess classification accuracy of the proposed methods.”